Strong coupling between scales in a multi-scalar model of urban dynamics

### 2021-04-12 Docs/Papers/Multiscale/paper/paper.tex
submission_ijgi
revision_multiscale

\documentclass[sustainability,article,submit,moreauthors,pdftex]{Definitions/mdpi}

@@

The modeling of urban growth and more generally the dynamics of urban systems is central to the design of sustainable territorial policies, through the understanding of past urbanisation processes and the anticipation of future urban trajectories. The design of sustainable future cities requires an historical knowledge of how past cities came to be and evolved \cite{batty2018inventing}. Several models have been proposed at different scales and integrating different dimensions of urban systems, such as models of land-use change at a mesoscopic scale or systems of cities models at a macroscopic scale \cite{pumain2017urban}.

@@

This paper contributes to the open question of multi-scalar models of urban dynamics by introducing a new simulation model which integrates a strong coupling between the mesoscopic scale and the macroscopic scale. The dynamics within each scale influence the other and reciprocally in an iterative way. More precisely, the model is simple in its components as we focus on the spatial structure of processes rather than on their multi-dimensionality. Therefore, we take into account only population variables, but both at the macroscopic scale of the system of cities and at the mesoscopic scale of the metropolitan area with an urban morphogenesis model. Our contribution is novel regarding previous works in particular regarding the following points: (i) the stylised model explicitly couples distinct scales and ontologies in a strong manner, most models operating only a weak coupling between scales (i.e. no reciprocal and dynamical feedbacks); (ii) the behavior of the model is systematically studied on synthetic systems of cities using model exploration methods.

@@

The main characteristic of our model is a strong coupling between the macroscopic scale of a system of cities with the mesoscopic scale of metropolitan areas. We consider simple urban dimensions with population variables only to describe cities, and use stylised processes. The system of cities evolves following a spatial interaction model as described by \cite{raimbault2020indirect}, assuming that interaction flows between cities will increase their attractivity and thus their population growth rate. Population updates are not done directly following migration flows as in stochastic urban models such as in \cite{james2018zipf}, but with the hypothesis of hidden variables (generally scaling with population) capturing for example economic dimensions, which influence population growth rate. At the mesoscopic scale, we consider the spatial distribution of population within an urban area and focus on the growth of urban form. To achieve this, the reaction-diffusion model of urban morphogenesis described by \cite{raimbault2018calibration} allows capturing concentration forces and dispersion forces, and has been shown to reproduce a large range of existing urban forms. Furthermore, \cite{raimbault2020comparison} showed that this model had a good coverage of the morphological space when compared to other models of urban morphogenesis.

To couple the scales, we assume that (i) macroscopic performance of a city will influence the choices made by planers in terms of land-use, and thus parameters of the morphogenesis mode locally - empirical evidence of such links have been suggested in the literature \cite{joy2015toronto} and we stylise them here; (ii) urban form influences the inner workings of a city through congestion and the fostering of exchanges between agents (individuals, firms), and finally the global insertion of the city - thus spatial interaction parameters are updated as a function of local urban form. This last link is the most discutable in terms of empirical evidence and implies many other dimensions than population only. We however work on a stylised explicative model which must keep a simple structure but on which a systematic knowledge of model behavior can be developed. We show in Fig.~\ref{fig:fig1} a schematic description of model structure.

\begin{figure}
%  figure: more details?
 \includegraphics[width=\linewidth]{figures/multiscale_morph.pdf}
 \caption{Schematic representation of model architecture. Urban areas with their own urban form are inserted into the macroscopic spatial interaction model, and strong coupling is achieved through top-down and bottom-up feedbacks.\label{fig:fig1}}
\end{figure}

@@

We consider $N$ urban areas, represented at the macroscopic scale by their population $P_j(t)$ at time $t$, and at the mesoscopic scale by a population grid $p_{kl}^{(j)}(t)$. The model runs for a total number $t_f$ of time steps, and we will assume that $\Delta t = 1$ for the sake of simplicity (the formulas can be generalized for arbitrary values of the time step, for example when running on real data with irregular time sampling).

The simulation model operates iteratively, and at each time step goes sequentially through the following steps:
\begin{enumerate}
	\item populations of urban areas are evolved at the macroscopic scale with the spatial interaction model (Equations \ref{eq:eq1} to \ref{eq:eq3} below);
	\item for each urban area, parameters of the moprhogenesis model are updated given parameters capturing typical scenarios: transit-oriented development or sprawl for the diffusion parameter, metropolization or uniformization for the aggregation parameter (Equations \ref{eq:eq4} to \ref{eq:eq6});
	\item local urban form are evolved conditionally to a fixed population increase and following the mescoscopic parameters;
	\item congestions and interactions within areas are synthesised following a cost function, which is used to update the macroscopic interaction parameters (Equations \ref{eq:eq7} and \ref{eq:eq8}).
\end{enumerate}

@@

\begin{table}
	\caption{Summary of model parameters\label{tab:parameters}}
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
		Type & Parameter & Process & Range \\\hline
		\multirow{4}{*}{Macro} & $g_i = g_0$ & Endogenous growth & $\left[0 ; 0.5\right]$ \\
		& $w_i = w_G$ & Interactions weight & $\left[0 ; 0.1\right]$ \\
		& $\gamma_i = \gamma_G$ & Interactions hierarchy & $\left[0 ; 5.0\right]$ \\
		& $d_i$ & Interactions decay & $\left[0 ; 1000\right]$ \\ \hline
		\multirow{4}{*}{Meso} & $\alpha_i$ & Aggregation & $\left[0 ; 5\right]$ \\
		& $\beta_i$ & Diffusion & $\left[0 ; 0.5\right]$ \\
		& $t_m$ & Urban growth speed & $\left[0 ; 20\right]$ \\
		& $n_d$ & Diffusion & $\left[0 ; 5\right]$ \\ \hline
		\multirow{4}{*}{Multiscale} & $\delta\alpha$ & Downward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\delta\beta$ & Downward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\delta d$ & Upward feedback & $\left[-0.5 ; 0.5\right]$ \\\hline
		& $\lambda$ & Cost of congestion & $\left[0 ; 2\right]$ \\\hline
	\end{tabular}
\end{table}

@@

@@

%%%
% ! fig placement

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{figures/allindics_hist.png}
	\caption{Statistical distribution of indicators for a sample of 10 parameter values.\label{fig:fig2}}
\end{figure}

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{figures/onefactor_allindics_mesoMacroDecayUpdateMax_errorbars.png}
	\caption{Variations of indicators as a function of $\delta d$, when all other parameters are fixed to their nominal value.\label{fig:fig3}}
\end{figure}

@@

@@

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{figures/deltaHierarchymacroPopulations-mesoTimeSteps_colorMacroInteractionDecay_facetmesoMacroDecayUpdateMax-macroMesoBetaUpdateMax.png}
	\caption{Macroscopic hierarchy of populations as a function of mesoscopic time steps $t_m$, for different values of interaction decay $d_G$ (colour), and of feedback parameters $\delta d$ (columns) and $\delta \beta$ (rows).\label{fig:fig4}}
\end{figure}

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{figures/deltamesoMorans-macroMesoAlphaUpdateMax_colormacroMesoBetaUpdateMax_facetmesoMacroCongestionCost-mesoMacroDecayUpdateMax_mesoBeta0_11.png}
	\caption{Relative evolution of average mescoscopic Moran, as a function of feedback parameter $\delta \beta$, for different values of feedback parameter $\delta \alpha$ (colour), and of feedback parameters $\delta d$ (columns) and $\lambda$ (rows).\label{fig:fig5}}
\end{figure}

\begin{figure}[h!]
	\includegraphics[width=\textwidth]{figures/deltamesoSlopes-macroMesoAlphaUpdateMax_colormacroMesoBetaUpdateMax_facetmesoMacroCongestionCost-mesoMacroDecayUpdateMax_mesoBeta0_11.png}
	\caption{Relative evolution of average mescoscopic population hierarchy, as a function of feedback parameter $\delta \beta$, for different values of feedback parameter $\delta \alpha$ (colour), and of feedback parameters $\delta d$ (columns) and $\lambda$ (rows).\label{fig:fig6}}
\end{figure}

\begin{figure}[h!]
	\begin{center}
	\includegraphics[width=0.6\linewidth]{figures/pareto_macroHierarchy-mesoHierarchy.png}
	\end{center}
	\caption{Pareto front obtained with a NSGA2 genetic algorithm, to minimize simultaneously hierarchy at the macroscopic and mesoscopic scales. Point size gives the number of stochastic samples and colour the value of $\delta \beta$.\label{fig:fig7}}
\end{figure}

@@

@@

@@

@@

>>>>

\documentclass[ijgi,article,submit,moreauthors,pdftex]{Definitions/mdpi}

@@

The modeling of urban growth and more generally the dynamics of urban systems is central to the design of sustainable territorial policies, through the understanding of past urbanisation processes and the anticipation of future urban trajectories. The design of sustainable future cities requires an historical knowledge of how past cities came to be and evolved \cite{batty2018inventing}. Several models have been developed at different scales and integrating different urban systems dimensions. Some typical examples are models of land-use change at a mesoscopic scale \cite{iacono2008models}, or systems of cities models at a macroscopic scale \cite{pumain2017urban}.

@@

This paper contributes to the open question of multi-scalar models of urban dynamics by introducing a new simulation model that integrates a strong coupling between the mesoscopic scale and the macroscopic scale. The dynamics within each scale influence the other and reciprocally in an iterative way. More precisely, the model is simple in its components as we focus on the spatial structure of processes rather than on their multi-dimensionality. Therefore, we take into account only population variables, but both at the macroscopic scale of the system of cities and at the mesoscopic scale of the metropolitan area with an urban morphogenesis model. Our contribution is novel regarding previous works in particular regarding the following points: (i) the stylised model explicitly couples distinct scales and ontologies in a strong manner, most models operating only a weak coupling between scales (i.e. no reciprocal and dynamical feedbacks); (ii) the behavior of the model is systematically studied on synthetic systems of cities using model exploration methods.

@@

The main characteristic of our model is a strong coupling between the macroscopic scale of a system of cities with the mesoscopic scale of metropolitan areas. We consider simple urban dimensions with population variables only to describe cities, and use stylised processes. The system of cities evolves following a spatial interaction model as described by \cite{raimbault2020indirect}, assuming that interaction flows between cities will increase their attractivity and thus their population growth rate. Population updates are not done directly following migration flows as in stochastic urban models such as in \cite{james2018zipf}, but with the hypothesis of hidden variables (generally scaling with population) capturing for example economic dimensions, which influence population growth rate. At the mesoscopic scale, we consider the spatial distribution of population within an urban area and focus on the growth of urban form. To achieve this, the reaction-diffusion model of urban morphogenesis described by \cite{raimbault2018calibration} allows capturing concentration forces and dispersion forces, and has been shown to reproduce a large range of existing urban forms. Furthermore, \cite{raimbault2020comparison} showed that this model had good coverage of the morphological space when compared to other models of urban morphogenesis.

To couple the scales, we assume that (i) macroscopic performance of a city will influence the choices made by planers in terms of land-use, and thus parameters of the morphogenesis mode locally - empirical evidence of such links have been suggested in the literature \cite{joy2015toronto} and we stylise them here; (ii) urban form influences the inner workings of a city through congestion and the fostering of exchanges between agents (individuals, firms), and finally the global insertion of the city - thus spatial interaction parameters are updated as a function of local urban form. This last link is the most % discutable % TODO slippery?
 in terms of empirical evidence and implies many other dimensions than population only. We however work on a stylised explicative model which must keep a simple structure but on which a systematic knowledge of model behavior can be developed. We show in Fig.~\ref{fig:fig1} a schematic description of model structure.

\begin{figure}[H]
%  figure: more details?
 \includegraphics[width=\linewidth]{figures/multiscale_morph.pdf}
 \caption{Schematic representation of model architecture. Urban areas with their own urban form are inserted into the macroscopic spatial interaction model, and strong coupling is achieved through top-down and bottom-up feedbacks.\label{fig:fig1}}
\end{figure}

@@

We consider $N$ urban areas, represented at the macroscopic scale by their population $P_j(t)$ at time $t$, and at the mesoscopic scale by a population grid $p_{kl}^{(j)}(t)$. The model runs for a total number $t_f$ of time steps, and we will assume that $\Delta t = 1$ for the sake of simplicity (the formulas can be generalised for arbitrary values of the time step, for example when running on real data with irregular time sampling).

The simulation model operates iteratively, and at each time step goes sequentially through the following steps:
\begin{enumerate}
	\item populations of urban areas are evolved at the macroscopic scale with the spatial interaction model (Equations \ref{eq:eq1} to \ref{eq:eq3} below);
	\item for each urban area, parameters of the moprhogenesis model are updated given parameters capturing typical scenarios: transit-oriented development or sprawl for the diffusion parameter, metropolisation or uniformisation for the aggregation parameter (Equations \ref{eq:eq4} to \ref{eq:eq6});
	\item local urban form are evolved conditionally to a fixed population increase and following the mescoscopic parameters;
	\item congestions and interactions within areas are synthesised following a cost function, which is used to update the macroscopic interaction parameters (Equations \ref{eq:eq7} and \ref{eq:eq8}).
\end{enumerate}

@@

@@

\begin{table}[H]
	\caption{Summary of model parameters\label{tab:parameters}}
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
		Type & Parameter & Process & Range \\\hline
		\multirow{4}{*}{Macro} & $g_i = g_0$ & Endogenous growth & $\left[0 ; 0.5\right]$ \\
		& $w_i = w_G$ & Interactions weight & $\left[0 ; 0.1\right]$ \\
		& $\gamma_i = \gamma_G$ & Interactions hierarchy & $\left[0 ; 5.0\right]$ \\
		& $d_i$ & Interactions decay & $\left[0 ; 1000\right]$ \\ \hline
		\multirow{4}{*}{Meso} & $\alpha_i$ & Aggregation & $\left[0 ; 5\right]$ \\
		& $\beta_i$ & Diffusion & $\left[0 ; 0.5\right]$ \\
		& $t_m$ & Urban growth speed & $\left[0 ; 20\right]$ \\
		& $n_d$ & Diffusion & $\left[0 ; 5\right]$ \\ \hline
		\multirow{4}{*}{Multiscale} & $\delta\alpha$ & Downward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\delta\beta$ & Downward feedback & $\left[-0.5 ; 0.5\right]$ \\
		& $\delta d$ & Upward feedback & $\left[-0.5 ; 0.5\right]$ \\\hline
		& $\lambda$ & Cost of congestion & $\left[0 ; 2\right]$ \\\hline
	\end{tabular}
\end{table}

@@

@@

\begin{figure}[H]
	\includegraphics[width=\linewidth]{figures/allindics_hist.png}
	\caption{Statistical distribution of indicators for a sample of 10 parameter values.\label{fig:fig2}}
\end{figure}

@@

@@

\begin{figure}[H]
	\includegraphics[width=\linewidth]{figures/deltaHierarchymacroPopulations-mesoTimeSteps_colorMacroInteractionDecay_facetmesoMacroDecayUpdateMax-macroMesoBetaUpdateMax.png}
	\caption{Macroscopic hierarchy of populations as a function of mesoscopic time steps $t_m$, for different values of interaction decay $d_G$ (colour), and of feedback parameters $\delta d$ (columns) and $\delta \beta$ (rows).\label{fig:fig4}}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=\linewidth]{figures/onefactor_allindics_mesoMacroDecayUpdateMax_errorbars.png}
	\caption{Variations of indicators as a function of $\delta d$, when all other parameters are fixed to their nominal value.\label{fig:fig3}}
\end{figure}

\begin{figure}[H]
	\includegraphics[width=\linewidth]{figures/deltamesoMorans-macroMesoAlphaUpdateMax_colormacroMesoBetaUpdateMax_facetmesoMacroCongestionCost-mesoMacroDecayUpdateMax_mesoBeta0_11.png}
	\caption{Relative evolution of average mescoscopic Moran, as a function of feedback parameter $\delta \beta$, for different values of feedback parameter $\delta \alpha$ (colour), and of feedback parameters $\delta d$ (columns) and $\lambda$ (rows).\label{fig:fig5}}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=\textwidth]{figures/deltamesoSlopes-macroMesoAlphaUpdateMax_colormacroMesoBetaUpdateMax_facetmesoMacroCongestionCost-mesoMacroDecayUpdateMax_mesoBeta0_11.png}
	\caption{Relative evolution of average mescoscopic population hierarchy, as a function of feedback parameter $\delta \beta$, for different values of feedback parameter $\delta \alpha$ (colour), and of feedback parameters $\delta d$ (columns) and $\lambda$ (rows).\label{fig:fig6}}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\linewidth]{figures/pareto_macroHierarchy-mesoHierarchy.png}
	\end{center}
	\caption{Pareto front obtained with a NSGA2 genetic algorithm, to minimize simultaneously hierarchy at the macroscopic and mesoscopic scales. Point size gives the number of stochastic samples and colour the value of $\delta \beta$.\label{fig:fig7}}
\end{figure}
